---
output: word_document
---
#Binomial Distribution#

The binomial distribution is a important distribution in the probability theory and statistics, it gives the the number of successes *k* in a sequence of *n independent Yes/No* experiments  with  **p**  as the probability of success of each yield

Lets understand the distribution  
If *p* is the probability that an event will happen in a single trial.  
*q =1-p* is the probability that it will fail to happen, 

then the probability that the event will happen exactly *k* in *n* trial is given
  
p(k) = $\frac{(n!)}{(n-k)! k!}(p^k)(1-p)^{(n-k)}$  where, k = 0, 1, 2, ..., n, 

$\frac{n!}{(n-k)! k!}$  is the binomial coefficient, hence the name binomial distributions. 

Lets understand this distribution more using some simple examples  

**Example 1** :   
Flip a fair coin 8 times and Probability of getting Tail X number of times.

**solution:**   
Outcome from first flip can be a head or tail, so total out come 2 outcomes;  Probability of a fair coin = 0.5.  .  

Using Binomial Distribution formula ,$\frac{(n!}{(n-k)! k!}(p^k)(1-p)^{(n-k)}$   
The probability of getting *0* tails in 8 flips =$\frac{(8!}{(8-0)! 0!}(0.5^0)(1-0.5)^{(8-0)}$= 1/256,

In similar ways,  
Probability of getting 0 tail in 8 flips = $\frac{(8!}{(8-0)! 0!}(0.5^0)(1-0.5)^{(8-0)}$  = `r (factorial(8)/(factorial(8-0) * factorial(0))) * (0.5^0) *(1-0.5)^(8-0) `   
Probability of getting 1 tail in 8 flips = $\frac{(8!}{(8-1)! 1!}(0.5^1)(1-0.5)^{(8-1)}$ = `r (factorial(8)/(factorial(8-1) * factorial(1))) * (0.5^1) *(1-0.5)^(8-1)`  
Probability of getting 2 tail in 8 flips  = $\frac{(8!}{(8-2)! 2!}(0.5^2)(1-0.5)^{(8-2)}$  = `r (factorial(8)/(factorial(8-2) * factorial(2))) * (0.5^2) *(1-0.5)^(8-2)`   
Probability of getting 3  tail in 8 flips = $\frac{(8!}{(8-3)! 3!}(0.5^3)(1-0.5)^{(8-3)}$  = `r (factorial(8)/(factorial(8-3) * factorial(3))) * (0.5^3) *(1-0.5)^(8-3)`   
Probability of getting 4  tail in 8 flips = $\frac{(8!}{(8-4)! 4!}(0.5^4)(1-0.5)^{(8-4)}$  = `r (factorial(8)/(factorial(8-4) * factorial(4))) * (0.5^4) *(1-0.5)^(8-4)`   
Probability of getting 5  tail in 8 flips = $\frac{(8!}{(8-5)! 5!}(0.5^5)(1-0.5)^{(8-5)}$  = `r (factorial(8)/(factorial(8-5) * factorial(5))) *(0.5^5) *(1-0.5)^(8-5)`  
Probability of getting 6  tail in 8 flips = $\frac{(8!}{(8-6)! 6!}(0.5^6)(1-0.5)^{(8-6)}$ = `r (factorial(8)/(factorial(8-6) * factorial(6))) * (0.5^6) *(1-0.5)^(8-6)`   
Probability of getting 7  tail in 8 flips = $\frac{(8!}{(8-7)! 7!}(0.5^7)(1-0.5)^{(8-7)}$  = `r (factorial(8)/(factorial(8-7) * factorial(7))) *  (0.5^7) *(1-0.5)^(8-7) `   
Probability of getting 8  tail in 8 flips = $\frac{(8!}{(8-8)! 8!}(0.5^8)(1-0.5)^{(8-8)}$  = `r (factorial(8)/(factorial(8-8) * factorial(8))) * (0.5^6) *(1-0.5)^(8-6)`    

The Binomial distribution is very important, when the sample size is huge, like flipping a coin 1000 times or more. Lets plot the curves and compare when the number of trial increase from 8 trials to 1000 trials. You can use the r syntax below
 
```{r}

n=8                                       # sample size i.e. 8 flips
k <- seq(0,8, by = 1)                     # Number of success 0, 1.....8
y <- c(dbinom(k, size = n ,prob = 0.5))   #dbinom gives the binomial dist.
par(mfrow=c(1,2))     
barplot( y,k, width= 1,space =0,
         main ="8 trials" , xlab = "Number of success" ,
         ylab="probability")
lines(k, (dbinom(k, n,0.5)), col = "red", lwd = 2)
abline(v=4, lty=2, col ="blue")

# 30 flips
n= 30
k <- seq(0,30, by = 1) 
y <- c(dbinom(k, size = n ,prob = 0.5))
barplot( y,k, width= 1,space =0, 
         main ="1000 trials", xlab = "Number of success" ,
         ylab="probability")
lines(k, (dbinom(k, n,0.5)), col = "red", lwd = 2)

```

**Example 2**  
If each gender has an independent 50% probability for each birth.What's the probability of getting 3 or more girl in 7 trials ?

probability of getting a girl *p* = 0.5  
probability of not getting a girl *q* = 1- 0.5= 0.5

*p(3)* = $\frac{(7!}{(7-3)! 3!}(0.5^3)(1-0.5)^{(7-3)}$ =  `r (factorial(7)/(factorial(3) * factorial(7-3))) *(0.5^3) *(0.5^(7-3))`  
*p(4)* = $\frac{(7!}{(7-4)! 4!}(0.5^3)(1-0.5)^{(7-4)}$ =  `r (factorial(7) / (factorial(4) * factorial(7-4))) *(0.5^4) *(0.5^(7-4))`   
*p(5)*= $\frac{(7!}{(7-5)! 5!}(0.5^5)(1-0.5)^{(7-5)}$ =  `r (factorial(7)/(factorial(5) * factorial(7-5))) *(0.5^5) *(0.5^(7-5))`  
*p(6)*= $\frac{(7!}{(7-6)! 6!}(0.5^3)(1-0.5)^{(7-6)}$ =  `r (factorial(7) / (factorial(6) * factorial(7-6))) *(0.5^6) *(0.5^(7-6))`  
*p(7)* = $\frac{(7!}{(7-7)! 7!}(0.5^7)(1-0.5)^{(7-7)}$ =  `r (factorial(7) / (factorial(7) * factorial(7-7))) *(0.5^7) *(0.5^(7-7))`

**The probability of 3 or more girl **(sum up above probabilities) = 
`r (factorial(7)/(factorial(3) * factorial(7-3))) *(0.5^3) *(0.5^(7-3)) + (factorial(7) / (factorial(4) * factorial(7-4))) *(0.5^4) *(0.5^(7-4)) + (factorial(7)/(factorial(5) * factorial(7-5))) *(0.5^5) *(0.5^(7-5)) +  (factorial(7) / (factorial(6) * factorial(7-6))) *(0.5^6) *(0.5^(7-6)) + 
(factorial(7) / (factorial(7) * factorial(7-7))) *(0.5^7) *(0.5^(7-7)) `

We can use dbinom or pdinom to get the probability of 3 or more girls: 
dbinom(3, size =7,prob =0.5) +  
dbinom(4, size =7,prob =0.5) +  
dbinom(5, size =7,prob =0.5) +  
dbinom(6, size =7,prob =0.5) +  
dbinom(7, size =7,prob =0.5) = `r dbinom(3, size =7,prob =0.5) +dbinom(4, size =7,prob =0.5) + dbinom(5, size =7,prob =0.5) + dbinom(6, size =7,prob =0.5) + dbinom(7, size =7,prob =0.5) `  
OR  
 pbinom(2, size = 7, prob = 0.5, lower.tail = FALSE) =`r  pbinom(2, size = 7, prob = 0.5, lower.tail = FALSE)`.
 
 Similarly,  

* The probability of getting 3 or less girls  dbinom(3, size =7,prob =0.5) +
dbinom(2, size =7,prob =0.5) + 
dbinom(1, size =7,prob =0.5) + 
dbinom(0, size =7,prob =0.5) = 
`r dbinom(3, size =7,prob =0.5) + dbinom(2, size =7,prob =0.5) + dbinom(1, size =7,prob =0.5) + dbinom(0, size =7,prob =0.5)`.  
* using **pbinom**, pbinom(3,size = 7, prob = 0.5, lower.tail =TRUE) = `r pbinom(3,size = 7, prob = 0.5, lower.tail =TRUE) `  

* The probability of getting 1 girl or less = **d**binom(0, size =7,prob =0.5) + dbinom(1, size =7,prob =0.5)* = `r dbinom(0, size =7,prob =0.5) + dbinom(1, size =7,prob =0.5) `  

* The probability of getting 1 or less girls =**p**binom(1 ,size = 7, prob = 0.5, lower.tail = TRUE)*  = `r pbinom(1 ,size = 7, prob = 0.5, lower.tail = TRUE)` 

* The probability of getting 1 or more girls=  *pbinom(0,size = 7, prob = 0.5, lower.tail =FALSE)* = `r pbinom(0,size = 7, prob = 0.5, lower.tail = FALSE)` 
 
 Lets see graphically  
```{r}
n=7
k <- seq(0,7, by = 1)        # Number of success 
y <- c(dbinom(k, size = n ,prob = 0.5))
barplot( y,k, width= 1,
         space =0,col = "blue",
         xlab = "Number of girl child" , ylab="probability")
```

**Example 3**
Suppose there are 10 multiple choice questions in an Math quiz. Each question has 4 possible answers, and only one of them is correct. Find the probability of having 5 or less correct answers if a student attempts to answer every question at random.

**Solution**  
There are 4 option, but only one correct answer , so probability = 1/4 = 0.25  
Total size = 10 
```{r}
n=11
k <- seq(0,11, by = 1) # Number of success 0, 1.....8
y <- c(dbinom(k, size = n ,prob = 0.25))
barplot( y,k, width= 1,space =0,
         xlab = "Success of getting K times Tail" , ylab="probability")
```  


The probability of  0 correct answer = dbinom(0,size =10, prob = 1/4) = `r dbinom(0,size =10, prob = 0.25) `  

The probability that , only 5 or less correct answer  = pbinom(5, size = 10, prob =0.25,lower.tail = TRUE) = `r pbinom(5, size = 10, prob = 0.25,lower.tail = TRUE) `  

The probability, Six or more correct answers =pbinom(5, size = 10, prob =0.25.lower.tail= FALSE) = `r pbinom(5, size = 10, prob = 0.25, lower.tail= FALSE) `
 
